---
title: "BiomasAnalysisV3"
author: "Gabriela Lopez and Tim Baker"
date: "Thursday, February 19, 2015"
output: html_document
---
**Requirements for using the ForestPlots.net Report template** 

(This section can be deleted once you're ready to publish your report)

This template is an R Markdown document that uses data from ForestPlots.net to create a user-friendly report about the latest results from your plots.

It produces an html or pdf document that reports above-ground biomass (AGB), stem density, wood mass density estimates and the frequency and number of species by family for your selected plots. It will also map the location of your plots, and visualise the rates of biomass change among the plots.

**Software requirements**

You will need to install R (version 3.1.2 or later) and RStudio <http://www.rstudio.com/> in your computer to modify and run the template.

In addition to a suite of packages that you will be asked to install when you first run an R Markdown file, you will also need to download the BiomasaFP package for estimating biomass and other forest parameters from data downloaded from ForestPlots.net. This can be obtained from <http://www.ForestPlots.net>.

Data requirements

To use this template, you need to download three data files from ForestPlots.net based on your plots: 

a) The Individuals csv file from the Advanced Search, 
b) The Wood density file from the query library 
c) The Plot View metadata file from the query library. 

The three files should correspond to the same Plot Views and should be saved in the same working directory, together with the BiomasaFP package.

**Instructions**

Once you have installed the programmes and downloaded your data, this template should be ready to run. Clicking the **Knit HTML** button will generate the document. To create a pdf document, select the **Knit PDF** option. To create a pdf file you will in addition need to have a TeX/LaTeX system on your computer;  you can download a suitable system from <http://www.miktex.org>; select the complete, rather than the basic system.

The text in blue below and in italics in the final document can be updated according to the dataset you're using. Indeed, if you are familiar with R, you can customize any aspect of the template to suit the output that you would like to obtain. 

For more details on using R Markdown see <http://rmarkdown.rstudio.com>. 

**Introduction**

This document summarizes Forest Inventory Data collected by the *RAINFOR network (www.rainfor.org)*.  The plots (plotviews) in this report conform to the following criteria: *multiple census plots*, *minimum diameter of 100mm*, *plot area greater than 0.2 ha*.

This document reports: a) Aboveground biomass (AGB), b) Basal Area (BA), c) wood density and d) frequency and number of species by Family for each Plot View.

-----
**Data processing**

The data was downloaded from ForestPlots.net (Lopez-Gonzalez et al., 2011; Lopez-Gonzalez et al., 2009) on *date*. The plots included in this report can be found in the table below. 

Note: click on a column header to sort the rows.

**Table 1. Plots included in this report**

```{r dataretrieving, cache=TRUE, results='asis', echo=FALSE}
#reading amd merging datasets
library (BiomasaFP)

# reads, cleans and merges the Census data file, metadata file and wd file.
mergedCensus <- mergefp("TestDataset.csv","Metadata10nov.csv","wd10nov.csv" )
Table1 <- SummaryPlotviews(mergedCensus)

```



```{r setOptions, message=FALSE, echo=FALSE}
library(googleVis)
op <- options(gvis.plot.tag='chart')

```


```{r ,results='asis', echo=FALSE}
Tbl1<- gvisTable (Table1)
plot(Tbl1, options= list (page='enable',height='automatic', width='automatic'))

```


The plot location is shown in the map in Figure 1.

**Figure 1. Map of Plot Location**

```{r, results='asis', echo=FALSE}
Table1$LatLong<-  paste(Table1$LatitudeDecimal,Table1$LongitudeDecimal, sep=":")
M <- gvisMap(Table1, locationvar = 'LatLong',  options= list (enableScrollWheel=TRUE, useMapTypeControl=TRUE, width =600, height=400, dataMode="markers"))
plot(M)

```

**Results**
Table 2 shows the Aboveground Biomass estimated using Chave (2005) Moist forest equation with heigt and wood density as parameters.Height was estimated using regional parameters  for a Weibull model (Feldpaush, 2011). 

**Table 2 Aboveground Biomass by PlotViewID and Census**
```{r, fig.height=4, results='asis',echo=FALSE, fig.path='figure/'}
library(BiomasaFP)
Table2 <- SummaryAGB(mergedCensus, AGBChv05MH)

Tbl2 <- gvisTable (Table2,options= list (page='enable',height='automatic', width='automatic'))
plot(Tbl2)
```


Aboveground biomass change was estimated as the difference in AGB between the first and last census, divided by the time elapsed (Figure 3). In Figure 3 biomass change is indicated by colour and the size of the marker represents the AGB weighted mean.

**Figure 3. Map of Change**

Note: This type of map includes an option for selecting region. In the configuration options of the following you can find the region codes: link https://developers.google.com/chart/interactive/docs/gallery/geomap
```{r, results='asis', echo=FALSE}
library(BiomasaFP)
Table3 <- AGBch(mergedCensus,AGBChv05MH)
Table3$AGBChangeyr <- Table3$AGBChange/(Table3$Last.Census.Mean.Date-Table3$Census.Mean.Date)
Table4 <- MeanWtAGB(mergedCensus,AGBChv05MH)
Table4a <- Table4[, c('PlotViewID', 'AGBWeightedMean')]
Table5 <- merge (Table3, Table4a, by = "PlotViewID")


suppressPackageStartupMessages(library(googleVis))
op <- options(gvis.plot.tag='chart')

Table5$LatLong<-  paste(Table5$LatitudeDecimal,Table5$LongitudeDecimal, sep=":")
M <- gvisGeoChart(Table5, locationvar = 'LatLong', sizevar='AGBWeightedMean', colorvar='AGBChangeyr' ,  options= list (enableScrollWheel=TRUE, useMapTypeControl=TRUE, width =600, height=400, region="005", dataMode="markers"))
plot(M)
```

**References**

Chave J, Coomes DA, Jansen S, Lewis SL, Swenson NG, Zanne AE. 2009. Towards a worldwide wood economics spectrum. Ecology Letters 12(4): 351-366. http://dx.doi.org/10.1111/j.1461-0248.2009.01285.x

Chave C, Andalo S, Brown, et al. 2005. Tree allometry and improved estimation of carbon stocks and balance in tropical forests. Oecologia 145 (1):87-99. doi:10.1007/s00442-005-0100-x.

Chave J, Rejou-Mechain M, Burquez A et al. 2014. Improved allometric models to estimate the aboveground biomass of tropical trees. Global Change Biology 20: 3177-3190. doi: 10.1111/gcb.12629

Feldpausch TR, Banin L, Phillips OL, Baker TR, Lewis SL et al. 2011. Height-diameter allometry of tropical forest trees. Biogeosciences 8 (5):1081-1106. doi:10.5194/bg-8-1081-2011.

Lopez-Gonzalez  G. (2014). BiomasaFP: Estimates Biomass for data dowloaded from ForestPlots.net. R package version 1.0.

Lopez-Gonzalez, G., Lewis, S.L., Burkitt, M. and Phillips, O.L. (2011). ForestPlots.net: a web application and research tool to manage and analyse tropical forest plot data. Journal of Vegetation Science 22: 610–613. doi: 10.1111/j.1654-1103.2011.01312.x

Lopez-Gonzalez, G., Lewis, S.L., Burkitt, M., Baker T.R. and Phillips, O.L. (2009). ForestPlots.net Database. www.forestplots.net. Date of extraction *[dd,mm,yy]*.

R Core Team (2014). R: A language and environment for statistical computing. R Foundation for Statistical Computing, Vienna, Austria. URL http://www.R-project.org/.

Zanne AE, Lopez-Gonzalez G, Coomes DA, Ilic J, Jansen S, Lewis SL, Miller RB, Swenson NG, Wiemann MC, Chave J. 2009. Data from: Towards a worldwide wood economics spectrum. Dryad Digital Repository. http://dx.doi.org/10.5061/dryad.234
